---
title: "STUBBS_Skills_Test_3"
author: "Josh Stubbs"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(skimr)
library(janitor)
library(tidyr)
library(broom)
library(tidyverse)
library(readxl)
library(lubridate)
library(gganimate)
library(transformr)
library(stringr)
library(gganimate)
library(gifski)
library(dplyr)
library(sf)
library(gplots)
library(scales)
library(easystats)
library(ggplot2)
library(readr)
```

# Goal of test
The goal is to us Rstudio to clean and analyze the 2 provided datasets. This will include cleaning data, modeling, and data analysis. 

# Task 1: "FacultySalaries_1995.csv"
The first task involves loading and cleaning the "FacultySalaries_1995.csv" dataset. It is important to understand what is in the data set to better analyze and clean it. 


# Read Data:

The code creates a file path and reads in a CSV file named "FacultySalaries_1995.csv" into a data frame. The read.csv function is used to read the CSV file.
```{r}

FacultySalaries_1995 <- read_csv("FacultySalaries_1995.csv")
fsalary <- FacultySalaries_1995


```


# Clean Data:

Pivot_longer column is used to stretch longer and split columns. 

```{r}

fsalary <- fsalary %>% 
  pivot_longer(cols = c("AvgFullProfSalary", "AvgAssocProfSalary", "AvgAssistProfSalary"),
               names_to = "Rank",
               values_to = "Salary") %>% 
  pivot_longer(cols = c("AvgFullProfComp", "AvgAssocProfComp", "AvgAssistProfComp"),
               names_to = "Rank_Comp",
               values_to = "Compensation") %>% 
  pivot_longer(cols = c("NumFullProfs", "NumAssocProfs", "NumAssistProfs", "NumInstructors"),
               names_to = "Rank_Num",
               values_to = "Staff_Number") %>% 
  mutate(Rank = case_when(Rank == "AvgFullProfSalary" ~ "Full",
                          Rank == "AvgAssocProfSalary" ~ "Assoc",
                          Rank == "AvgAssistProfSalary" ~ "Assist"))


```

# Filter Data:

clears rows where the "Tier" variable is equal to "VIIB".

```{r}

fsalary <- fsalary[!(fsalary$Tier=="VIIB"), ]
```

# Visual Data:

create a boxplot of salary distribution (Salary) across different faculty ranks (Rank), with each rank having a different fill color.


```{r pt.2, warning=FALSE}

fsalary %>% 
  ggplot(aes(x=Rank,y=Salary)) +
  geom_boxplot(aes(fill=Rank)) +
  facet_wrap(~Tier) +
  theme_minimal() +
  theme(legend.key.size = unit(1, "cm"),
        legend.title = element_text(size = 19, color = "black"),
        legend.text = element_text(size = 13, color = "black"),
        axis.text.x = element_text(angle = 65, size = 13, color = "black"),
        axis.text.y = element_text(size = 13, color = "black"),
        axis.title.x = element_text(size = 19, color = "black"),
        axis.title.y = element_text(size = 19, color = "black"),
        strip.text.x = element_text(size = 13, color = "black"))

```

# Export Figure:

The ggsave function is then used to save the created plot as a JPEG image.

```{r warning=FALSE}
ggsave("./STUBBS_fig_1.jpg")
```

# Task 2: ANOVA Modeling
In the second task an (ANOVA) model is made to examine the influence of "State," "Tier," and "Rank" on "Salary."

# Build and Analyze ANOVA Model:

This code creates a generalized linear model (GLM) where Salary is shown as a function  of Rank, State, and Tier in the data set. The summary function of the fitted GLM model, shows coefficients, standard errors, t-values, and p-values.

```{r}
model_1 <- glm(data = fsalary, 
            formula = Salary ~ Rank + State + Tier )
summary(model_1)
```

# Export ANOVA Summary:

The sink function saves the output to a file ("Salary_ANOVA_Summary.txt") instead of the console, providing a written summary of the ANOVA results.

```{r pt.3,}

anova(model_1)
summary(model_1)

sink("./Salary_ANOVA_Summary.txt") 
anova(model_1)
sink(NULL)

```


# Task 3: Juniper_Oils.csv

The task is to clean and prepare data for analysis. This dataset contains measurements of chemical concentrations in the essential oil of cedar trees over various years since a burn event.

# Clean and Tidy Another Dataset:
This code reads the data from "Juniper_Oils.csv" and tidies it by converting chemical concentrations from wide to long format using pivot_longer.

```{r pt.4,}

Juniper_Oils <- read_csv("Juniper_Oils.csv")

Juniper_Oils <- Juniper_Oils %>% 
  pivot_longer(cols = c("alpha-pinene","para-cymene","alpha-terpineol","cedr-9-ene",
                        "alpha-cedrene","beta-cedrene","cis-thujopsene","alpha-himachalene",
                        "beta-chamigrene","cuparene","compound 1","alpha-chamigrene",
                        "widdrol","cedrol","beta-acorenol","alpha-acorenol","gamma-eudesmol",
                        "beta-eudesmol","alpha-eudesmol","cedr-8-en-13-ol","cedr-8-en-15-ol",
                        "compound 2","thujopsenal"),
               names_to = "ChemicalID",
               values_to = "Concentration")
```

# Task 4: Graphical Representation
Make graph of relationship between "YearsSinceBurn" and "Concentration" for different chemical compounds. 

# Data Prep.:
Use ggplot to make and plot information. 

```{r}
Juniper_Oils %>% 
  ggplot(aes(x=YearsSinceBurn, y=Concentration)) +
  facet_wrap(~ChemicalID, scales = "free_y") +
  geom_smooth(method = loess) +
  theme_minimal() +
  labs(x = "Years Since Burn",
       title = "Concentration by Chemical") +
  theme(plot.title = element_text(hjust = 0.5))


```

# Exporting the Plot:

ggsave saves the plot as an image file named "STUBBS_Fig_2.jpg" in the current working directory.
```{r}
# export to millerberg_Fig_2.jpg
ggsave("./STUBBS_Fig_2.jpg")
```

## Task 5: Generalized Linear Model
The final task is to use a (GLM) and tidy() function. 

# Generalized Linear Model

```{r}
model2 <- glm(data = Juniper_Oils, formula = Concentration ~ ChemicalID * YearsSinceBurn)

model2_sum <- summary(Juniper_Oils)

```

# Tidy Data Frame:
Tidy function from the broom package is used to tidy up the model results.

```{r}
tidy <- tidy(model2)

```

# Subset Tidy Data Frame:
This give p-value and prints it for viewing. 
```{r}
ttss <- tidy[c(tidy$p.value <= 0.05), ]
print(ttss)
```
